---
title: "data_analysis_script"
author: "Peter Czingraber"
date: "2026-04-21"
output: pdf_document
---

### Installing packages

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
```

### Importing data

```{r}
raw_pilot_data = read_csv("raw_data.csv")
```

### Data cleaning

#### Flagging practice trials

```{r}
raw_pilot_data = raw_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(block_marker = coalesce(str_detect(stimulus, regex("Blokk\\s*1\\s*kezdődik", ignore_case = TRUE)),
                                 FALSE),
         start_seen = cummax(trial_index == 0),
         marker_seen = cummax(block_marker),
         trial_exp = if_else(start_seen == 1 & marker_seen == 0, "practice", "experimental")) %>% 
  ungroup()
```

#### Flagging previous trials

```{r}
raw_pilot_data = raw_pilot_data %>%
  mutate(prev_congruency = lag(congruency, n = 4),
         prev_correct = lag(correct, n = 4),
         correct = ifelse(correct == "TRUE", 1, 0),
         prev_correct = ifelse(prev_correct == "TRUE", 1, 0),
         first_trial = ifelse(lag(str_detect(stimulus, "Blokk \\d+ kezdődik"), 4) == TRUE, 1, 0),
         rt = as.numeric(rt))
```

### Filtering subjects with accuracy below 60%

```{r}
accuracy = raw_pilot_data %>% 
      group_by(subj_code) %>%
      summarize(all_accuracy = mean(correct, na.rm = TRUE))

raw_pilot_data = raw_pilot_data %>% 
  select(-any_of("all_accuracy")) %>% 
  left_join(accuracy, by = "subj_code") %>%
  filter(all_accuracy > 0.60, first_trial != 1) 
```

#### Filtering data

```{r}
filtered_pilot_data = raw_pilot_data %>%
  filter(correct == 1,
         prev_correct == 1,
         rt >= 150 | is.na(rt))
```

```{r}
processed_pilot_data = filtered_pilot_data %>% 
  select(
    -stimulus,
    -response,
    -trial_type,
    -trial_index,
    -plugin_version,
    -time_elapsed,
    -success,
    -question_order,
    -correct_response,
    -block_marker,
    -marker_seen,
    -prev_correct,
    -start_seen,
    -trial_exp,
    -first_trial,
    -name)
```

### Descriptive analysis function

```{r}
dstats = function(x, na.omit=FALSE){
  if (na.omit)
  x = x[!is.na(x)]
  min = min(x)
  max = max(x)
  m = mean(x)
  median = median(x)
  n = length(x)
  stds = sd(x)
  skew = 3*(m-median)/stds
  kurt = (length(x) * sum((x - mean(x))^4)) / (sum((x - mean(x))^2)^2) - 3

  
#Return all calculated values to vectors
  return(c(
    min = min,
    max = max,
    mean = m,
    median = median,
    sd = stds,
    skewness = skew,
    kurtosis = kurt
  ))
}
```

### Descriptive statistics for aggregated RT values

```{r}
mean_rt_df = processed_pilot_data %>% 
  group_by(subj_code) %>% 
  summarise(mean_rt_congruent = mean(rt[congruency == "congruent"]),
            mean_rt_incongruent = mean(rt[congruency == "incongruent"]))

myvars = c("mean_rt_congruent", "mean_rt_incongruent")
sapply(mean_rt_df[myvars], dstats)
```

### Density plot

```{r}
density_rt = ggplot(processed_pilot_data, aes(rt))
density_rt + geom_density()
```

### Violinplots for CE

#### Appropriate data structure for visualization

```{r}
ce_rt = processed_pilot_data %>% 
  group_by(subj_code, congruency) %>% 
  summarise(mean_rt = mean(rt))
```

```{r}
violin = ggplot(ce_rt, aes(congruency, mean_rt, fill = congruency)) +
  geom_violin(trim = TRUE, bounds = c(0, Inf), width = 0.5) +
  geom_boxplot(width = 0.1, colour = c("black", "white"), outlier.shape = NA) +
  labs(title = "Congruency Effect", x = "Congruency", y = "Mean reaction time (ms)") +
  papaja::theme_apa(base_family = "Times") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("congruent" = "#FAC775", "incongruent" = "#BA7517"))
#ggsave("ce_violin.png", plot = violin, width = 16, height = 9, units = "cm", dpi = 320)
print(violin)
```

### Assumption check for H1

```{r}
assumption_h1 = with(mean_rt_df, mean_rt_incongruent - mean_rt_congruent)
shapiro.test(assumption_h1)
```

#### If the assumption of normality is not met, we will use Wilcoxon signed-rank test for H1

```{r}
wilcox.test(mean_rt_df$mean_rt_incongruent, mean_rt_df$mean_rt_congruent, paired = TRUE, alternative = "greater")
```

### H1 - Congruency Effect (paired samples t-test)

```{r}
t.test(mean_rt_df$mean_rt_incongruent, mean_rt_df$mean_rt_congruent, paired = T, alternative = "greater")
```



















